Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-11-27 Moltres.tex
Address monolithic vs segregated solution question.

>>>>

\subsection{Monolithic vs. Segregated Solution}

Relative performance of monolithic vs. segregated solution methods for large
systems of equations is an active area of research. Monolithic solvers are
generally regarded to be more robust than their segregated counterparts but are
often perceived as being too expensive for large-scale problems. Work conducted
by Heil et. al. \cite{heil2008solvers} demonstrates that monolithic solvers are
competitive with segregated techniques for weakly coupled physics; moreover, for
strongly coupled problems the segregated methods struggle to achieve
convergence. It is noted by the authors that use of monolithic solvers
on large 3D problems requires efficient preconditioning to be effective.

Monolithic and segregated solution techniques were compared in Moltres using the
2D-axisymmetric problem but with delayed neutron precursors removed. In this
case starting from arbitrary initial conditions the neutron multiplication
exhibits damped oscillation around the critical state. Feedback between the
power deposition by neutrons and temperature modification of neutron macroscopic
group constants make this a very tightly coupled problem. Using adaptive time
stepping based on the number of non-linear iterations at each time step,
steady-state neutron fluxes and temperature profiles are achieved in 358 time
steps and in 4 minutes of compute time on a single core using full-coupling
(e.g. neutron fluxes and temperature in the same matrix). Alternatively,
a solution was also attempted splitting neutron and temperature solves and
performing Picard iterations until convergence. Ultimately, a steady-state
solution could not be achieved using the segregated method. After 670
time-steps, the Picard solve hits the maximum number of allowed iterations (100)
after only halving the neutronics residual. Before reaching the fail-point, the
segregated solution required numerous Picard iterations to converge (68, 37, 22
for the previous three time-steps for example), resulting in a total compute
time of 37 minutes before failure. These results suggest that for solving tight coupling
between neutronics and temperature, a monololithic solve approach is more effective.

### 2017-11-27 revise/revise.tex
Address monolithic vs segregated solution question.

\question It's somehow interesting that ``For some three-dimensional
simulations, the number of elements in the mesh and total number of degrees of
freedom exceed one million and ten million respectively. To handle problems of
this size, we ran Moltres on up to 608 cores.'' Normally problems of few million
degrees of freedom can be handled with few tens of cores. The authors should
discuss more extensively the performances of their code in order to give the
reader the possibility to compare them with those of other tools. In particular,
it is possible that slow performances may derive from trying to achieve a fully
coupled solution (i.e., all the equations in the same matrix), which seems to be
the case in this work. In this sense and to better support the claims of the
paper, it would be important to compare the performances of a segregated and a
fully coupled solver, both in terms of accuracy and computational time. In
addition to helping supporting the claims of the paper, this would represent a
significant contribution to the scientific community, since many of the
currently available tools employ segregated solvers, and since the advantages
and disadvantages of a fully coupled solver are often a subject of discussion.
\begin{solution}
 <++>
\end{solution}

>>>>

\question It's somehow interesting that ``For some three-dimensional
simulations, the number of elements in the mesh and total number of degrees of
freedom exceed one million and ten million respectively. To handle problems of
this size, we ran Moltres on up to 608 cores.'' Normally problems of few million
degrees of freedom can be handled with few tens of cores. The authors should
discuss more extensively the performances of their code in order to give the
reader the possibility to compare them with those of other tools. In particular,
it is possible that slow performances may derive from trying to achieve a fully
coupled solution (i.e., all the equations in the same matrix), which seems to be
the case in this work. In this sense and to better support the claims of the
paper, it would be important to compare the performances of a segregated and a
fully coupled solver, both in terms of accuracy and computational time. In
addition to helping supporting the claims of the paper, this would represent a
significant contribution to the scientific community, since many of the
currently available tools employ segregated solvers, and since the advantages
and disadvantages of a fully coupled solver are often a subject of discussion.
\begin{solution}
        It is definitely true that for many problems an operator-splitting or
        segregated solution method will be faster and require less
        computational resources. A fully-coupled Newton's method is
        computationally expensive. We could in theory use fewer cores to solve
        our million DOF problems and have the problem run longer. However,
        particularly in our department and with our collaboration with Illinois's
        National Center for Supercomputing Applications, we are encouraged to
        use the computational resources at our disposal, so we elect to run our
        large problems as quickly as possible on Blue Waters. Additionally, the
        developers of PetSc recommend using approximately 20k degrees of freedom
        per process, which is a rule of thumb we generally adhere to.

        For coupling between neutrons and temperature, we performed a comparison
        between monolothic and segregated solution techniques. The results are
        outlined in a new results subsection placed immediately before the
        Conclusion section. In brief, for this problem which exhibits such strong
        feedback between physics, we were unable to achieve convergence using the
        segregated method.
\end{solution}
